---
title: "PM566 Final Project"
author: "Songru Wang"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---
This is my PM566 Final Project website.


Introdution:
Since nowadays there are lot of smokers around the world and many policies have been applied to curb the number of smokers, I would like to figure out whether the number of smokers is actually decreased or not. This data is about "Never smoked trend for 1995-2010". I've found the data from Centers for Disease Control and Prevention website. 

Then I want to figure out whether non-smokers in each different city become less or not as time goes by, and I want to find the relationship between the change in the number of smokers and the state.


Description of the data:
While analysing the original data table,I found there are 876 obervations, and 7 variables. In this data set, 56 states (including several islands next to the US) are included. And there is no missing values in the data table.  
```{r,cache=TRUE}
X2013_08_01T20_31_13_000Z <- read_csv("2013-08-01T20_31_13.000Z.csv")
smoke<- X2013_08_01T20_31_13_000Z
dim(smoke)
head(smoke)
tail(smoke)
unique(smoke$State)
```

```{r,cache=TRUE}
str(smoke)
table(smoke$Year)
table(smoke$State)
any(is.na(smoke$`Smoke everyday`))
any(is.na(smoke$`Smoke some days`))
any(is.na(smoke$`Former smoker`))
any(is.na(smoke$`Never smoked`))
```


Data Analsis
```{r,cache=TRUE}
library(skimr)
skim(smoke)
# Find the relationship between year and the average number of each category
smoke_avg<-
  smoke%>%
  group_by(Year)%>%
  summarize(
    S_everyday_avg=mean(`Smoke everyday`),
    S_someday_avg=mean(`Smoke some days`),
    S_former_avg=mean(`Former smoker`),
    S_never_avg=mean(`Never smoked`)
  )
smoke_avg
```

```{r,cache=TRUE}
smoke_avg%>%
  ggplot(mapping = aes(x=Year,y=S_everyday_avg))+
  geom_point()+
  geom_smooth(method=lm,col="black")

smoke_avg%>%
  ggplot(mapping = aes(x=Year,y=S_someday_avg))+
  geom_point()+
  geom_smooth(method=lm,col="black")

smoke_avg%>%
  ggplot(mapping = aes(x=Year,y=S_former_avg))+
  geom_point()+
  geom_smooth(method=lm,col="black")

smoke_avg%>%
  ggplot(mapping = aes(x=Year,y=S_never_avg))+
  geom_point()+
  geom_smooth(method=lm,col="black")
```

According to the four graphs, I found the number of people who smoke everyday shows a decrease as times goes by. The number of people who smoke someday shows a gradual increase as times goes by.The number of people who never smoke becomes more as time goes by.The number of people who is former smokers, in other words, the people who quit smoking, becomes more as times goes by.



Then I divide these 56 states into 5 geographical categories: Northeast, Southwest, West, Southeast, and Midwest.

Alaska is hard to define which region it belongs to, so I leave it as NA value.Guam is an organized, unincorporated territory of the United States, so I leave it as NA value.District of Columbia is not a state.
I also choose to not define the geographic region of Virgin Island and Hawaii

Smoke Everyday in each geographic region shows a decreasing tendency. Smoke Somedays in each geographical region shows a slight increase. Former Smoker showed decreasing tendency in the West region, increasing tendency in Northeast, Midwest, and Southeast


```{r,cache=TRUE}
smoke<-
  smoke%>%
  mutate(Geo_cate=case_when(
    State == "Connecticut" ~ "Northeast",
    State == "Maine" ~ "Northeast",
    State == "Massachusetts" ~ "Northeast",
    State == "New Hampshire" ~ "Northeast",
    State == "Rhode Island" ~ "Northeast",
    State == "Vermont" ~ "Northeast",
    State == "New Jersey" ~ "Northeast",
    State == "New York" ~ "Northeast",
    State == "Delaware" ~ "Northeast",
    State == "Pennsylvania" ~ "Northeast",
    State == "Alabama" ~ "Southeast",
    State == "Arkansas" ~ "Southeast",
    State == "Florida" ~ "Southeast",
    State == "Georgia" ~ "Southeast",
    State == "Kentucky" ~ "Southeast",
    State == "Louisiana" ~ "Southeast",
    State == "Mississippi" ~ "Southeast",
    State == "North Carolina" ~ "Southeast",
    State == "South Carolina" ~ "Southeast",
    State == "Tennessee" ~ "Southeast",
    State == "Virginia" ~ "Southeast",
    State == "West Virginia" ~ "Southeast",
    State == "Arizona" ~ "Southwest",
    State == "Colorado" ~ "Southwest",
    State == "Utah" ~ "Southwest",
    State == "Nevada" ~ "Southwest",
    State == "New Mexico" ~ "Southwest",
    State == "Idaho" ~ "West",
    State == "Montana" ~ "West",
    State == "Wyoming" ~ "West",
    State == "California" ~ "West",
    State == "Washington" ~ "West",
    State == "Oregon" ~ "West",
    State == "Hawaii" ~ "West",
    State == "Oklahoma" ~ "Southwest",
    State == "Texas" ~ "Southwest",
    State == "Illinois" ~ "Midwest",
    State == "Indiana" ~ "Midwest",
    State == "Iowa" ~ "Midwest",
    State == "Kansas" ~ "Midwest",
    State == "Michigan" ~ "Midwest",
    State == "Minnesota" ~ "Midwest",
    State == "Missouri" ~ "Midwest",
    State == "Nebraska" ~ "Midwest",
    State == "North Dakota" ~ "Midwest",
    State == "Ohio" ~ "Midwest",
    State == "South Dakota" ~ "Midwest",
    State == "Wisconsin" ~ "Midwest",
  ))
```

```{r,cache=TRUE}

# Smoke Everyday 
smoke%>%
  filter(Geo_cate=='Northeast')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke everyday`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='Southwest')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke everyday`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='West')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke everyday`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='Southeast')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke everyday`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='Midwest')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke everyday`))+
  geom_point()+
  geom_smooth(method = lm,col='black')


# Smoke Somedays
smoke%>%
  filter(Geo_cate=='Northeast')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke some days`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='Southwest')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke some days`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='West')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke some days`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='Southeast')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke some days`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='Midwest')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke some days`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

#Former Smoker
smoke%>%
  filter(Geo_cate=='Northeast')%>%
  ggplot(mapping = aes(x=Year,y=`Former smoker`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='Southwest')%>%
  ggplot(mapping = aes(x=Year,y=`Former smoker`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='West')%>%
  ggplot(mapping = aes(x=Year,y=`Former smoker`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='Southeast')%>%
  ggplot(mapping = aes(x=Year,y=`Former smoker`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='Midwest')%>%
  ggplot(mapping = aes(x=Year,y=`Former smoker`))+
  geom_point()+
  geom_smooth(method = lm,col='black')
```


Distribution on the map
a. "Smoking Everyday"
```{r,cache=TRUE}
library(leaflet)

commu.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=smoke$`Smoke everyday`)
leaflet(smoke1)%>%
  addProviderTiles('CartoDB.VoyagerLabelsUnder')%>%
  addCircles(
    lat = ~lat,lng = ~lon,
    label = ~paste0(round(`Smoke everyday`,2)),color = ~commu.pal(`Smoke everyday`),
    opacity = 1, fillOpacity = 1, radius = 500
  )%>%
  addLegend('bottomleft',pal = commu.pal,values = smoke1$`Smoke everyday`,title = 'The number of people smoking everyday',opacity = 1)
```


b. "Smoke Some Days"
```{r,cache=TRUE}
library(leaflet)

commu.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=smoke1$`Smoke some days`)
leaflet(smoke1)%>%
  addProviderTiles('CartoDB.VoyagerLabelsUnder')%>%
  addCircles(
    lat = ~lon,lng = ~lat,
    label = ~paste0(round(`Smoke some days`,2)),color = ~commu.pal(`Smoke some days`),
    opacity = 1, fillOpacity = 1, radius = 500
  )%>%
  addLegend('bottomleft',pal = commu.pal,values = smoke1$`Smoke some days`,title = 'The number of people smoking some days',opacity = 1)
```


c. "Former Smoker"
```{r,cache=TRUE}
library(leaflet)

commu.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=smoke1$`Former smoker`)
leaflet(smoke1)%>%
  addProviderTiles('CartoDB.VoyagerLabelsUnder')%>%
  addCircles(
    lat = ~lon,lng = ~lat,
    label = ~paste0(round(`Former smoker`,2)),color = ~commu.pal(`Former smoker`),
    opacity = 1, fillOpacity = 1, radius = 500
  )%>%
  addLegend('bottomleft',pal = commu.pal,values = smoke1$`Former smoker`,title = 'The number of people who is former smoker',opacity = 1)
```


D. Never Smoke
```{r,cache=TRUE}
library(leaflet)

commu.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=smoke1$`Never smoked`)
leaflet(smoke1)%>%
  addProviderTiles('CartoDB.VoyagerLabelsUnder')%>%
  addCircles(
    lat = ~lon,lng = ~lat,
    label = ~paste0(round(`Never smoked`,2)),color = ~commu.pal(`Never smoked`),
    opacity = 1, fillOpacity = 1, radius = 500
  )%>%
  addLegend('bottomleft',pal = commu.pal,values = smoke1$`Never smoked`,title = 'The number of people who never smoke',opacity = 1)
```

Conclusion:
Based on the analysis, we can conclude that the number of smokers decreases as time goes by, the number of former smokers increases, and the number of people who never smoke increases. This is possibly due to education, policy, or the taxes on cigerattes. However, the changes in different geographical regions is different, which needs further data collection and data analysis.

