---
title: "Smoking-Y&N"
author: "Songru Wang"
data: "Never smoked trend for 1995-2010"
date: "10/19/2021"
output:
  github_document:
  html_document:
    html_preview: false
always_allow_html: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
library(dtplyr)
library(dplyr)
options(repos = c(CRAN = "http://cran.rstudio.com"))
```



# This data is about "Never smoked trend for 1995-2010". I've found the data from Centers for Disease Control and Prevention website. Then I want to figure out whether non-smokers in each different city become less or not as time goes by, and I want to find the relationship between the change in the number of smokers and the state.


1. Import Data
```{r,cache=TRUE}
X2013_08_01T20_31_13_000Z <- read_csv("2013-08-01T20_31_13.000Z.csv")
smoke<- X2013_08_01T20_31_13_000Z
```


Find dimensions
```{r,cache=TRUE}
dim(smoke)
head(smoke)
tail(smoke)
unique(smoke$State)
# I found there are 876 obervations, and 7 variables.
# In this data set, 56 states are included.
```

Take a look at the variable
```{r,cache=TRUE}
str(smoke)
table(smoke$Year)
table(smoke$State)
any(is.na(smoke$`Smoke everyday`))
any(is.na(smoke$`Smoke some days`))
any(is.na(smoke$`Former smoker`))
any(is.na(smoke$`Never smoked`))
# I found there is no NA value in these four columns
```

Take a closer look at the variable
```{r,cache=TRUE}
library(skimr)
skim(smoke)

# Find the relationship between year and the average number of each category
smoke_avg<-
  smoke%>%
  group_by(Year)%>%
  summarize(
    S_everyday_avg=mean(`Smoke everyday`),
    S_someday_avg=mean(`Smoke some days`),
    S_former_avg=mean(`Former smoker`),
    S_never_avg=mean(`Never smoked`)
  )
smoke_avg
```


2. Data visualization
first: Year&smoking status
```{r,cache=TRUE}
smoke_avg%>%
  ggplot(mapping = aes(x=Year,y=S_everyday_avg))+
  geom_point()+
  geom_smooth(method=lm,col="black")

smoke_avg%>%
  ggplot(mapping = aes(x=Year,y=S_someday_avg))+
  geom_point()+
  geom_smooth(method=lm,col="black")

smoke_avg%>%
  ggplot(mapping = aes(x=Year,y=S_former_avg))+
  geom_point()+
  geom_smooth(method=lm,col="black")

smoke_avg%>%
  ggplot(mapping = aes(x=Year,y=S_never_avg))+
  geom_point()+
  geom_smooth(method=lm,col="black")

## According to the four graphs, I found the number of people who smoke everyday shows a decrease as times goes by. 
#The number of people who smoke someday shows a gradual increase as times goes by.
#The number of people who never smoke becomes more as time goes by.
#The number of people who is former smokers, in other words, the people who quit smoking, becomes more as times goes by 

```



Let's divide these 56 states into 5 geographical categories: Northeast, Southwest, West, Southeast, and Midwest
```{r,cache=TRUE}
smoke<-
  smoke%>%
  mutate(Geo_cate=case_when(
    State == "Connecticut" ~ "Northeast",
    State == "Maine" ~ "Northeast",
    State == "Massachusetts" ~ "Northeast",
    State == "New Hampshire" ~ "Northeast",
    State == "Rhode Island" ~ "Northeast",
    State == "Vermont" ~ "Northeast",
    State == "New Jersey" ~ "Northeast",
    State == "New York" ~ "Northeast",
    State == "Delaware" ~ "Northeast",
    State == "Pennsylvania" ~ "Northeast",
    State == "Alabama" ~ "Southeast",
    State == "Arkansas" ~ "Southeast",
    State == "Florida" ~ "Southeast",
    State == "Georgia" ~ "Southeast",
    State == "Kentucky" ~ "Southeast",
    State == "Louisiana" ~ "Southeast",
    State == "Mississippi" ~ "Southeast",
    State == "North Carolina" ~ "Southeast",
    State == "South Carolina" ~ "Southeast",
    State == "Tennessee" ~ "Southeast",
    State == "Virginia" ~ "Southeast",
    State == "West Virginia" ~ "Southeast",
    State == "Arizona" ~ "Southwest",
    State == "Colorado" ~ "Southwest",
    State == "Utah" ~ "Southwest",
    State == "Nevada" ~ "Southwest",
    State == "New Mexico" ~ "Southwest",
    State == "Idaho" ~ "West",
    State == "Montana" ~ "West",
    State == "Wyoming" ~ "West",
    State == "California" ~ "West",
    State == "Washington" ~ "West",
    State == "Oregon" ~ "West",
    State == "Hawaii" ~ "West",
    State == "Oklahoma" ~ "Southwest",
    State == "Texas" ~ "Southwest",
    State == "Illinois" ~ "Midwest",
    State == "Indiana" ~ "Midwest",
    State == "Iowa" ~ "Midwest",
    State == "Kansas" ~ "Midwest",
    State == "Michigan" ~ "Midwest",
    State == "Minnesota" ~ "Midwest",
    State == "Missouri" ~ "Midwest",
    State == "Nebraska" ~ "Midwest",
    State == "North Dakota" ~ "Midwest",
    State == "Ohio" ~ "Midwest",
    State == "South Dakota" ~ "Midwest",
    State == "Wisconsin" ~ "Midwest",
  ))
# Alaska is hard to define which region it belongs to, so I leave it as NA value.
# Guam is an organized, unincorporated territory of the United States, so I leave it as NA value.
# District of Columbia is not a state.
# I also choose to not define the geographic region of Virgin Island and Hawaii.
```


Plot the smoking status by the factor of geographic regions
```{r,cache=TRUE}

# Smoke Everyday in each geographic region shows a decreasing tendency.
smoke%>%
  filter(Geo_cate=='Northeast')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke everyday`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='Southwest')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke everyday`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='West')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke everyday`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='Southeast')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke everyday`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='Midwest')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke everyday`))+
  geom_point()+
  geom_smooth(method = lm,col='black')


# Smoke Somedays in each geographical region shows a slight increase.
smoke%>%
  filter(Geo_cate=='Northeast')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke some days`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='Southwest')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke some days`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='West')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke some days`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='Southeast')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke some days`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='Midwest')%>%
  ggplot(mapping = aes(x=Year,y=`Smoke some days`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

#Former Smoker showed decreasing tendency in the West region, increasing tendency in Northeast, Midwest, and Southeast
smoke%>%
  filter(Geo_cate=='Northeast')%>%
  ggplot(mapping = aes(x=Year,y=`Former smoker`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='Southwest')%>%
  ggplot(mapping = aes(x=Year,y=`Former smoker`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='West')%>%
  ggplot(mapping = aes(x=Year,y=`Former smoker`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='Southeast')%>%
  ggplot(mapping = aes(x=Year,y=`Former smoker`))+
  geom_point()+
  geom_smooth(method = lm,col='black')

smoke%>%
  filter(Geo_cate=='Midwest')%>%
  ggplot(mapping = aes(x=Year,y=`Former smoker`))+
  geom_point()+
  geom_smooth(method = lm,col='black')
```


Using leaflet to see how number of smokers spread out in different cities
1. find the longtitude and latitude of each city
```{r,cache=TRUE}
dat <- c("Alabama (32.840569999605975, -86.63186000013877)",
"Alaska (64.84507999974238, -147.72205999986895)",
"Arizona (34.86596999961597, -111.76380999973156)",
"Arkansas (34.748649999697875, -92.27448999971358)",
"California (37.638300000444815, -120.99958999997835)",
"Colorado (38.842890000173554, -106.13314000041055)",
"Connecticut (41.56265999995918, -72.6498400002157)",
"Delaware (39.00883000020451, -75.57774000040052)",
"District of Columbia (38.89036999987576, -77.03195999965413)",
"Florida (28.932039999846268, -81.9289599999039)",
"Georgia (32.83967999993223, -83.62758000031658)",
"Hawaii (21.304850000427336, -157.85774999956269)",
"Idaho (43.682590000228515, -114.36368000023168)",
"Illinois (40.485010000411364, -88.99770999971656)",
"Indiana (39.76690999989677, -86.14996000035359)",
"Iowa (42.469390000048634, -93.81649000001335)",
"Kansas (38.34774000000118, -98.20077999969709)",
"Kentucky (37.645969999815804, -84.77496999996538)",
"Louisiana (31.31265999975932, -92.44567999993188)",
"Maine (45.25423000041434, -68.9850299999344)",
"Maryland (39.29057999976732, -76.6092600004485)",
"Massachusetts (42.27687000005062, -72.08269000004333)",
"Michigan (44.661320000317914, -84.71438999959867)",
"Minnesota (46.3556499998478, -94.79419999982997)",
"Mississippi (32.7455100000866, -89.53803000008429)",
"Missouri (38.63578999960896, -92.5663000000448)",
"Montana (47.06653000015956, -109.42441999998289)",
"Nebraska (41.6410400000961, -99.36572999973953)",
"Nevada (39.49323999972637, -117.07183999971608)",
"New Hampshire (43.65595000019255, -71.50036000041354)",
"New Jersey (40.13056999960594, -74.2736899996936)",
"New Mexico (34.52088000011207, -106.24057999976702)",
"New York (42.82699999955048, -75.54396999981549)",
"North Carolina (35.46624999963797, -79.1593199999179)",
"North Dakota (47.475320000018144, -100.11841999998285)",
"Ohio (40.06020999969189, -82.40426000019869)",
"Oklahoma (35.4720099999617, -97.52034999975251)",
"Oregon (44.567449999917756, -120.15502999983448)",
"Pennsylvania (40.79372999993973, -77.86069999960512)",
"Rhode Island (41.70828000002217, -71.5224700001902)",
"South Carolina (33.99855000018255, -81.0452500001872)",
"South Dakota (44.353130000049646, -100.37353000040906)",
"Tennessee (35.68094000038087, -85.77449000011325)",
"Texas (31.82724000022597, -99.42676999973554)",
"Utah (39.36070000030492, -111.58712999994941)",
"Vermont (43.625379999687425, -72.51764000028561)",
"Virginia (37.54268000028196, -78.45789000012326)",
"Washington (47.522280000022135, -120.47001000026114)",
"West Virginia (38.66550999958696, -80.71263999973604)",
"Wisconsin (44.39319000021851, -89.81636999977553)",
"Wyoming (43.23553999957147, -108.10982999975454)")
dat <- data.frame(state = dat, stringsAsFactors = FALSE)

dat_new <- data.frame(
  state = gsub("\\s*\\(.+", "", dat$state, perl = TRUE),
  lat   = stringr::str_extract(dat$state, "(?<=\\()[0-9.-]+"),
  lon   = stringr::str_extract(dat$state, "[0-9.-]+(?=\\))")
)

dat_new$lon <- as.numeric(dat_new$lon)
dat_new$lat <- as.numeric(dat_new$lat)

str(dat_new)

smoke1<-smoke
smoke1 <- left_join(smoke1,dat_new, 
              by = c("State" = "state"))
```



2. Using leaflet
a. "Smoking Everyday"
```{r,cache=TRUE}
library(leaflet)

commu.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=smoke$`Smoke everyday`)
leaflet(smoke1)%>%
  addProviderTiles('CartoDB.VoyagerLabelsUnder')%>%
  addCircles(
    lat = ~lat,lng = ~lon,
    label = ~paste0(round(`Smoke everyday`,2)),color = ~commu.pal(`Smoke everyday`),
    opacity = 1, fillOpacity = 1, radius = 500
  )%>%
  addLegend('bottomleft',pal = commu.pal,values = smoke1$`Smoke everyday`,title = 'The number of people smoking everyday',opacity = 1)
```


b. "Smoke Some Days"
```{r,cache=TRUE}
library(leaflet)

commu.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=smoke1$`Smoke some days`)
leaflet(smoke1)%>%
  addProviderTiles('CartoDB.VoyagerLabelsUnder')%>%
  addCircles(
    lat = ~lon,lng = ~lat,
    label = ~paste0(round(`Smoke some days`,2)),color = ~commu.pal(`Smoke some days`),
    opacity = 1, fillOpacity = 1, radius = 500
  )%>%
  addLegend('bottomleft',pal = commu.pal,values = smoke1$`Smoke some days`,title = 'The number of people smoking some days',opacity = 1)
```


c. "Former Smoker"
```{r,cache=TRUE}
library(leaflet)

commu.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=smoke1$`Former smoker`)
leaflet(smoke1)%>%
  addProviderTiles('CartoDB.VoyagerLabelsUnder')%>%
  addCircles(
    lat = ~lon,lng = ~lat,
    label = ~paste0(round(`Former smoker`,2)),color = ~commu.pal(`Former smoker`),
    opacity = 1, fillOpacity = 1, radius = 500
  )%>%
  addLegend('bottomleft',pal = commu.pal,values = smoke1$`Former smoker`,title = 'The number of people who is former smoker',opacity = 1)
```


D. Never Smoke
```{r,cache=TRUE}
library(leaflet)

commu.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=smoke1$`Never smoked`)
leaflet(smoke1)%>%
  addProviderTiles('CartoDB.VoyagerLabelsUnder')%>%
  addCircles(
    lat = ~lon,lng = ~lat,
    label = ~paste0(round(`Never smoked`,2)),color = ~commu.pal(`Never smoked`),
    opacity = 1, fillOpacity = 1, radius = 500
  )%>%
  addLegend('bottomleft',pal = commu.pal,values = smoke1$`Never smoked`,title = 'The number of people who never smoke',opacity = 1)
```


